\textbf{Effective application of machine learning to microbiome based classification problems}

### 2019-10-15 submission/manuscript.Rmd
edit typos and table S1

Machine learning (ML) modeling of the human microbiome has the potential to identify microbial biomarkers and aid in diagnosis of many diseases such as inflammatory bowel disease, diabetes, and colorectal cancer. Progress has been made towards developing ML models that predict health outcomes using bacterial abundances, but inconsistent adoption of training and evaluation methods call the validity of these models into question. Furthermore, there appears to be a preference by many researchers to favor increased model complexity over interpretability. To overcome these challenges, we trained seven models that used fecal 16S rRNA sequence data to predict the presence of colonic screen relevant neoplasias (SRNs; n=490 patients, 261 controls and 229 cases). We developed a reusable open-source pipeline to train, validate and interpret the models. To show the effect of model selection, we assessed the predictive performance, interpretability, and training time of L2-regularized logistic regression, L1 and L2-regularized support vector machines (SVM) with linear and radial basis function kernels, decision trees, random forest, and gradient boosted trees (XGBoost). The random forest model performed best at detecting SRNs with an AUROC of 0.695 [IQR 0.651-0.739]  but was slow to train (83.2 h) and not immidiately interpretable. Despite its simplicity, L2-regularized logistic regression followed random forest in predictive performance with an AUROC of 0.680 [IQR 0.625-0.735], trained faster (12 min), and yielded a model that was inherently interpretable. Our analysis highlights the importance of choosing an ML approach based on the goal of the study, as the choice will inform expectations of performance and interpretability.

@@

As the number of people represented in human microbiome datasets grow, there is an increasing desire to use microbiome data to diagnose disease. However, the structure of the human microbiome is remarkably variable among individuals to the point where it is often difficult to identify the bacterial populations that are associated with diseases using traditional statistical models. This variation is likely due to the ability of many bacterial populations to fill the same niche such that different populations cause the same disease in different individuals. Furthermore, a growing number of studies have shown that it is rare for a single bacterial species to be associated with a disease. Instead, subsets of the microbiome account for differences in health. Traditional statistical approaches do not adequately account for the variation in the human microbiome and typically consider the protective or risk effects of each bacterial population seperately [@segata_metagenomic_2011]. Recently, machine learning (ML) models have grown in popularity among microbiome researchers as our ability to to sample large numbers of individuals has grown; such models can effectively account for the interpersonal microbiome variation and the ecology of disease.

@@

Second, to analyze non-linear models we interpreted the feature importance using permutation importance [@10.1093/bioinformatics/btq134]. Whereas the absolute feature weights were determined from the trained models, here we measured importance using the held-out test data. Permutation importance analysis is a post hoc explanation of the model, in which we randomly permuted groups of perfectly correlated features together and other features individually across the two groups in the held-out test data. We then calculated how much the predictive performance of the model (i.e, testing AUROC values) decreased when each OTU or group of OTUs was randomly permuted. We ranked the OTUs based on how much the median testing AUROC decreased when it was permuted; the OTU with the largest decrease ranked highest [Figure 4]. Among the twenty OTUs with the largest impact, there was only one OTU (OTU 822) that was shared among all of the models; however, we found three OTUs (OTUs 58, 110, 367) that were important in each of the tree-based models. Similarly, the random forest and XGBoost models, shared four of the most important OTUs (OTUs 2, 12, 361, 477). Permutation analysis results also revealed that with the exception of the decision tree model, removal of any individual OTU had minimal impact on model performance. For example, if OTU 367 was permuted across the samples in the decision tree model, the median AUROC dropped from `r paste(round(median((all %>% filter(model=="Decision_Tree"))$test_aucs), 3))` to `r paste(round(DT$imp, 3))`. In contrast, if the same OTU was permuted in the random forest model, the AUROC only dropped from `r paste(round(median((all %>% filter(model=="Random_Forest"))$test_aucs), 3))` to `r paste(format(round(RF$imp, 3), nsmall=3))` which indicated high degree of collinearity in the dataset. Permutation analysis allowed us to gauge the importance of each OTU in non-linear models and partially account for collinearity by grouping correlated OTUs to determine their impact as a group.

@@

**Figure 2. Generalization and classification performance of ML models using AUROC values of all cross validation and testing performances.** The median AUROC for diagnosing individuals with SRN using bacterial abundances was higher than chance (depicted by horizontal line at 0.50) for all the ML models. The predictive performance of random forest model was higher than other ML models,s though not significantly (p=0.5). L2-regularized logistic regression, XGBoost, L2-regularized SVM with linear and radial basis function kernel performances were not significantly different from one another. The boxplot shows quartiles at the box ends and the median as the horizontal line in the box. The whiskers show the farthest points that were not outliers. Outliers were defined as those data points that are not within 1.5 times the interquartile ranges. 
\newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}

@@

\makecell[l]{Source \\ of data} & \makecell[l]{Input data are not appropriate \\ for intended application. \\ (e.g. input data can be \\ used for diagnosis \\ but not prognosis)} & \makecell[l]{Data are generated according \\ to study design and appropriate \\ for intended application} & \makecell[l]{Data are available \\ to predict intended \\ outcomes and will be \\ available in the future } \\ \hline

\makecell[l]{Study \\ cohort} & \makecell[l]{Test data resampled to \\ remove class imbalance} & \makecell[l]{Test data are reflective \\ of the population which \\ the model will eventually \\ be applied to} & \makecell[l]{Model tested on \\ multiple cohorts \\ with different \\ class balances} \\ \hline

\makecell[l]{Model \\ selection} & \makecell[l]{No justification for \\ classification method} & \makecell[l]{Model choice is justified \\ for intended application} & \makecell[l]{Different modeling \\ choices tested} \\ \hline

\makecell[l]{Model \\ development} & \makecell[l]{No hyperparameter tuning} & \makecell[l]{Tested across multiple \\ hyperparameter settings} & \makecell[l]{Cross-validation performed \\ on training set \\ with hyperparameter \\ grid search} \\ \hline

\makecell[l]{Model \\ evaluation} & \makecell[l]{Performance reported on the data \\ used to train the model} & \makecell[l]{Performance reported on \\ held-out test data} & \makecell[l]{Performance reported on \\ multiple held-out test data } \\ \hline

\makecell[l]{Evaluation \\ metrics} & \makecell[l]{Reported a single metric that is \\ not appropriate for intended \\ application  (e.g. when predicting \\ rare outcome, accuracy metric is \\ not reliable)} & \makecell[l]{Reported multiple metrics} & \makecell[l]{Reported multiple metrics \\ with confidence intervals} \\ \hline

\makecell[l]{Model \\ interpretation} & \makecell[l]{No model interpretation} & \makecell[l]{Follow-up analyses to \\ determine what is driving \\ model performance} & \makecell[l]{Hypotheses on important \\ features are generated \\ and tested} \\ \hline

\makecell[l]{Model \\ specification} & \makecell[l]{Only predictive performance \\ reported} & \makecell[l]{Model described \\ (e.g. feature weights)} & \makecell[l]{How to use the model \\ is explained} \\ \hline
\end{tabular}
\newpage

>>>>

Machine learning (ML) modeling of the human microbiome has the potential to identify microbial biomarkers and aid in diagnosis of many diseases such as inflammatory bowel disease, diabetes, and colorectal cancer. Progress has been made towards developing ML models that predict health outcomes using bacterial abundances, but inconsistent adoption of training and evaluation methods call the validity of these models into question. Furthermore, there appears to be a preference by many researchers to favor increased model complexity over interpretability. To overcome these challenges, we trained seven models that used fecal 16S rRNA sequence data to predict the presence of colonic screen relevant neoplasias (SRNs; n=490 patients, 261 controls and 229 cases). We developed a reusable open-source pipeline to train, validate and interpret the models. To show the effect of model selection, we assessed the predictive performance, interpretability, and training time of L2-regularized logistic regression, L1 and L2-regularized support vector machines (SVM) with linear and radial basis function kernels, decision trees, random forest, and gradient boosted trees (XGBoost). The random forest model performed best at detecting SRNs with an AUROC of 0.695 [IQR 0.651-0.739] but was slow to train (83.2 h) and not immediately interpretable. Despite its simplicity, L2-regularized logistic regression followed random forest in predictive performance with an AUROC of 0.680 [IQR 0.625-0.735], trained faster (12 min), and was inherently interpretable. Our analysis highlights the importance of choosing an ML approach based on the goal of the study, as the choice will inform expectations of performance and interpretability.

@@

As the number of people represented in human microbiome datasets grow, there is an increasing desire to use microbiome data to diagnose disease. However, the structure of the human microbiome is remarkably variable among individuals to the point where it is often difficult to identify the bacterial populations that are associated with diseases using traditional statistical models. This variation is likely due to the ability of many bacterial populations to fill the same niche such that different populations cause the same disease in different individuals. Furthermore, a growing number of studies have shown that it is rare for a single bacterial species to be associated with a disease. Instead, subsets of the microbiome account for differences in health. Traditional statistical approaches do not adequately account for the variation in the human microbiome and typically consider the protective or risk effects of each bacterial population separately [@segata_metagenomic_2011]. Recently, machine learning (ML) models have grown in popularity among microbiome researchers as our ability to sample large numbers of individuals has grown; such models can effectively account for the interpersonal microbiome variation and the ecology of disease.

@@

Second, to analyze non-linear models we interpreted the feature importance using permutation importance [@10.1093/bioinformatics/btq134]. Whereas the absolute feature weights were determined from the trained models, here we measured importance using the held-out test data. Permutation importance analysis is a post hoc explanation of the model, in which we randomly permuted groups of perfectly correlated features together and other features individually across the two groups in the held-out test data. We then calculated how much the predictive performance of the model (i.e., testing AUROC values) decreased when each OTU or group of OTUs was randomly permuted. We ranked the OTUs based on how much the median testing AUROC decreased when it was permuted; the OTU with the largest decrease ranked highest [Figure 4]. Among the twenty OTUs with the largest impact, there was only one OTU (OTU 822) that was shared among all of the models; however, we found three OTUs (OTUs 58, 110, 367) that were important in each of the tree-based models. Similarly, the random forest and XGBoost models, shared four of the most important OTUs (OTUs 2, 12, 361, 477). Permutation analysis results also revealed that with the exception of the decision tree model, removal of any individual OTU had minimal impact on model performance. For example, if OTU 367 was permuted across the samples in the decision tree model, the median AUROC dropped from `r paste(round(median((all %>% filter(model=="Decision_Tree"))$test_aucs), 3))` to `r paste(round(DT$imp, 3))`. In contrast, if the same OTU was permuted in the random forest model, the AUROC only dropped from `r paste(round(median((all %>% filter(model=="Random_Forest"))$test_aucs), 3))` to `r paste(format(round(RF$imp, 3), nsmall=3))` which indicated high degree of collinearity in the dataset. Permutation analysis allowed us to gauge the importance of each OTU in non-linear models and partially account for collinearity by grouping correlated OTUs to determine their impact as a group.

@@

**Figure 2. Generalization and classification performance of ML models using AUROC values of all cross validation and testing performances.** The median AUROC for diagnosing individuals with SRN using bacterial abundances was higher than chance (depicted by horizontal line at 0.50) for all the ML models. The predictive performance of random forest model was higher than other ML models, though not significantly (p=0.5). L2-regularized logistic regression, XGBoost, L2-regularized SVM with linear and radial basis function kernel performances were not significantly different from one another. The boxplot shows quartiles at the box ends and the median as the horizontal line in the box. The whiskers show the farthest points that were not outliers. Outliers were defined as those data points that are not within 1.5 times the interquartile ranges. 
\newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}

@@

\makecell[l]{Source \\ of data} & \makecell[l]{Input data are not appropriate \\ for intended application (e.g. input \\ data provided only from \\ patients with carcinomas but \\ model is expected to predict \\ advanced adenomas.)} & \makecell[l]{Input data are generated \\ according to study design and \\ they are appropriate for intended \\ application.} & \makecell[l]{Input data are available \\ to predict intended outcomes \\ and will be available \\ in the future (e.g. same OTU \\ assignments for new fecal \\ samples)} \\ \hline

\makecell[l]{Study \\ cohort} & \makecell[l]{Test data resampled to \\ remove class imbalance. \\ (e.g. test data only includes \\ patients with carcinomas but \\ model is trained to predict \\ healthy colons as well.)}  & \makecell[l]{Test data are reflective \\ of the population which \\ the model will eventually \\ be applied to.} & \makecell[l]{Model tested on multiple \\ cohorts with different \\ class balances.} \\ \hline

\makecell[l]{Model \\ selection} & \makecell[l]{No justification for \\ classification method.} & \makecell[l]{Model choice is justified \\ for intended application.} & \makecell[l]{Different modeling \\ choices tested.} \\ \hline

\makecell[l]{Model \\ development} & \makecell[l]{No hyperparameter tuning.} & \makecell[l]{Model is tested across multiple \\ hyperparameter settings.} & \makecell[l]{Hyperparameter grid search \\ performed by cross-validation \\ on the training set.} \\ \hline

\makecell[l]{Model \\ evaluation} & \makecell[l]{Performance reported on the \\data used to train the model.} & \makecell[l]{Performance reported on \\ held-out test data.} & \makecell[l]{Performance reported on \\ multiple held-out test data.} \\ \hline

\makecell[l]{Evaluation \\ metrics} & \makecell[l]{Reported a single performance \\ metric that is not appropriate for \\ intended application  (e.g. when \\ predicting rare outcome, \\ accuracy metric is not \\ reliable).} & \makecell[l]{Reported a single performance \\ metric that is appropriate \\ for intended application with \\ confidence intervals.} & \makecell[l]{Reported multiple metrics \\ with confidence intervals.} \\ \hline

\makecell[l]{Model \\ interpretation} & \makecell[l]{No model interpretation.} & \makecell[l]{Follow-up analyses to \\ determine what is driving \\ model performance.} & \makecell[l]{Hypotheses on important \\ features are generated \\ and tested.} \\ \hline

\makecell[l]{Model \\ specification} & \makecell[l]{Only predictive performance \\ reported.} & \makecell[l]{Model described \\ (e.g. feature weights).} & \makecell[l]{How to use the model \\ is explained.} \\ \hline
\end{tabular}
\newpage

### 2019-10-15 submission/manuscript.tex
edit typos and table S1

Machine learning (ML) modeling of the human microbiome has the potential
to identify microbial biomarkers and aid in diagnosis of many diseases
such as inflammatory bowel disease, diabetes, and colorectal cancer.
Progress has been made towards developing ML models that predict health
outcomes using bacterial abundances, but inconsistent adoption of
training and evaluation methods call the validity of these models into
question. Furthermore, there appears to be a preference by many
researchers to favor increased model complexity over interpretability.
To overcome these challenges, we trained seven models that used fecal
16S rRNA sequence data to predict the presence of colonic screen
relevant neoplasias (SRNs; n=490 patients, 261 controls and 229 cases).
We developed a reusable open-source pipeline to train, validate and
interpret the models. To show the effect of model selection, we assessed
the predictive performance, interpretability, and training time of
L2-regularized logistic regression, L1 and L2-regularized support vector
machines (SVM) with linear and radial basis function kernels, decision
trees, random forest, and gradient boosted trees (XGBoost). The random
forest model performed best at detecting SRNs with an AUROC of 0.695
{[}IQR 0.651-0.739{]} but was slow to train (83.2 h) and not immidiately
interpretable. Despite its simplicity, L2-regularized logistic
regression followed random forest in predictive performance with an
AUROC of 0.680 {[}IQR 0.625-0.735{]}, trained faster (12 min), and
yielded a model that was inherently interpretable. Our analysis
highlights the importance of choosing an ML approach based on the goal
of the study, as the choice will inform expectations of performance and
interpretability.

@@

As the number of people represented in human microbiome datasets grow,
there is an increasing desire to use microbiome data to diagnose
disease. However, the structure of the human microbiome is remarkably
variable among individuals to the point where it is often difficult to
identify the bacterial populations that are associated with diseases
using traditional statistical models. This variation is likely due to
the ability of many bacterial populations to fill the same niche such
that different populations cause the same disease in different
individuals. Furthermore, a growing number of studies have shown that it
is rare for a single bacterial species to be associated with a disease.
Instead, subsets of the microbiome account for differences in health.
Traditional statistical approaches do not adequately account for the
variation in the human microbiome and typically consider the protective
or risk effects of each bacterial population seperately (1). Recently,
machine learning (ML) models have grown in popularity among microbiome
researchers as our ability to to sample large numbers of individuals has
grown; such models can effectively account for the interpersonal
microbiome variation and the ecology of disease.

@@

Second, to analyze non-linear models we interpreted the feature
importance using permutation importance (36). Whereas the absolute
feature weights were determined from the trained models, here we
measured importance using the held-out test data. Permutation importance
analysis is a post hoc explanation of the model, in which we randomly
permuted groups of perfectly correlated features together and other
features individually across the two groups in the held-out test data.
We then calculated how much the predictive performance of the model
(i.e, testing AUROC values) decreased when each OTU or group of OTUs was
randomly permuted. We ranked the OTUs based on how much the median
testing AUROC decreased when it was permuted; the OTU with the largest
decrease ranked highest {[}Figure 4{]}. Among the twenty OTUs with the
largest impact, there was only one OTU (OTU 822) that was shared among
all of the models; however, we found three OTUs (OTUs 58, 110, 367) that
were important in each of the tree-based models. Similarly, the random
forest and XGBoost models, shared four of the most important OTUs (OTUs
2, 12, 361, 477). Permutation analysis results also revealed that with
the exception of the decision tree model, removal of any individual OTU
had minimal impact on model performance. For example, if OTU 367 was
permuted across the samples in the decision tree model, the median AUROC
dropped from 0.601 to 0.525. In contrast, if the same OTU was permuted
in the random forest model, the AUROC only dropped from 0.695 to 0.680
which indicated high degree of collinearity in the dataset. Permutation
analysis allowed us to gauge the importance of each OTU in non-linear
models and partially account for collinearity by grouping correlated
OTUs to determine their impact as a group.

@@

\textbf{Figure 2. Generalization and classification performance of ML
models using AUROC values of all cross validation and testing
performances.} The median AUROC for diagnosing individuals with SRN
using bacterial abundances was higher than chance (depicted by
horizontal line at 0.50) for all the ML models. The predictive
performance of random forest model was higher than other ML models,s
though not significantly (p=0.5). L2-regularized logistic regression,
XGBoost, L2-regularized SVM with linear and radial basis function kernel
performances were not significantly different from one another. The
boxplot shows quartiles at the box ends and the median as the horizontal
line in the box. The whiskers show the farthest points that were not
outliers. Outliers were defined as those data points that are not within
1.5 times the interquartile ranges. \newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}

@@

\makecell[l]{Source \\ of data} & \makecell[l]{Input data are not appropriate \\ for intended application. \\ (e.g. input data can be \\ used for diagnosis \\ but not prognosis)} & \makecell[l]{Data are generated according \\ to study design and appropriate \\ for intended application} & \makecell[l]{Data are available \\ to predict intended \\ outcomes and will be \\ available in the future } \\ \hline

\makecell[l]{Study \\ cohort} & \makecell[l]{Test data resampled to \\ remove class imbalance} & \makecell[l]{Test data are reflective \\ of the population which \\ the model will eventually \\ be applied to} & \makecell[l]{Model tested on \\ multiple cohorts \\ with different \\ class balances} \\ \hline

\makecell[l]{Model \\ selection} & \makecell[l]{No justification for \\ classification method} & \makecell[l]{Model choice is justified \\ for intended application} & \makecell[l]{Different modeling \\ choices tested} \\ \hline

\makecell[l]{Model \\ development} & \makecell[l]{No hyperparameter tuning} & \makecell[l]{Tested across multiple \\ hyperparameter settings} & \makecell[l]{Cross-validation performed \\ on training set \\ with hyperparameter \\ grid search} \\ \hline

\makecell[l]{Model \\ evaluation} & \makecell[l]{Performance reported on the data \\ used to train the model} & \makecell[l]{Performance reported on \\ held-out test data} & \makecell[l]{Performance reported on \\ multiple held-out test data } \\ \hline

\makecell[l]{Evaluation \\ metrics} & \makecell[l]{Reported a single metric that is \\ not appropriate for intended \\ application  (e.g. when predicting \\ rare outcome, accuracy metric is \\ not reliable)} & \makecell[l]{Reported multiple metrics} & \makecell[l]{Reported multiple metrics \\ with confidence intervals} \\ \hline

\makecell[l]{Model \\ interpretation} & \makecell[l]{No model interpretation} & \makecell[l]{Follow-up analyses to \\ determine what is driving \\ model performance} & \makecell[l]{Hypotheses on important \\ features are generated \\ and tested} \\ \hline

\makecell[l]{Model \\ specification} & \makecell[l]{Only predictive performance \\ reported} & \makecell[l]{Model described \\ (e.g. feature weights)} & \makecell[l]{How to use the model \\ is explained} \\ \hline
\end{tabular}\newpage

>>>>

Machine learning (ML) modeling of the human microbiome has the potential
to identify microbial biomarkers and aid in diagnosis of many diseases
such as inflammatory bowel disease, diabetes, and colorectal cancer.
Progress has been made towards developing ML models that predict health
outcomes using bacterial abundances, but inconsistent adoption of
training and evaluation methods call the validity of these models into
question. Furthermore, there appears to be a preference by many
researchers to favor increased model complexity over interpretability.
To overcome these challenges, we trained seven models that used fecal
16S rRNA sequence data to predict the presence of colonic screen
relevant neoplasias (SRNs; n=490 patients, 261 controls and 229 cases).
We developed a reusable open-source pipeline to train, validate and
interpret the models. To show the effect of model selection, we assessed
the predictive performance, interpretability, and training time of
L2-regularized logistic regression, L1 and L2-regularized support vector
machines (SVM) with linear and radial basis function kernels, decision
trees, random forest, and gradient boosted trees (XGBoost). The random
forest model performed best at detecting SRNs with an AUROC of 0.695
{[}IQR 0.651-0.739{]} but was slow to train (83.2 h) and not immediately
interpretable. Despite its simplicity, L2-regularized logistic
regression followed random forest in predictive performance with an
AUROC of 0.680 {[}IQR 0.625-0.735{]}, trained faster (12 min), and was
inherently interpretable. Our analysis highlights the importance of
choosing an ML approach based on the goal of the study, as the choice
will inform expectations of performance and interpretability.

@@

As the number of people represented in human microbiome datasets grow,
there is an increasing desire to use microbiome data to diagnose
disease. However, the structure of the human microbiome is remarkably
variable among individuals to the point where it is often difficult to
identify the bacterial populations that are associated with diseases
using traditional statistical models. This variation is likely due to
the ability of many bacterial populations to fill the same niche such
that different populations cause the same disease in different
individuals. Furthermore, a growing number of studies have shown that it
is rare for a single bacterial species to be associated with a disease.
Instead, subsets of the microbiome account for differences in health.
Traditional statistical approaches do not adequately account for the
variation in the human microbiome and typically consider the protective
or risk effects of each bacterial population separately (1). Recently,
machine learning (ML) models have grown in popularity among microbiome
researchers as our ability to sample large numbers of individuals has
grown; such models can effectively account for the interpersonal
microbiome variation and the ecology of disease.

@@

Second, to analyze non-linear models we interpreted the feature
importance using permutation importance (36). Whereas the absolute
feature weights were determined from the trained models, here we
measured importance using the held-out test data. Permutation importance
analysis is a post hoc explanation of the model, in which we randomly
permuted groups of perfectly correlated features together and other
features individually across the two groups in the held-out test data.
We then calculated how much the predictive performance of the model
(i.e., testing AUROC values) decreased when each OTU or group of OTUs
was randomly permuted. We ranked the OTUs based on how much the median
testing AUROC decreased when it was permuted; the OTU with the largest
decrease ranked highest {[}Figure 4{]}. Among the twenty OTUs with the
largest impact, there was only one OTU (OTU 822) that was shared among
all of the models; however, we found three OTUs (OTUs 58, 110, 367) that
were important in each of the tree-based models. Similarly, the random
forest and XGBoost models, shared four of the most important OTUs (OTUs
2, 12, 361, 477). Permutation analysis results also revealed that with
the exception of the decision tree model, removal of any individual OTU
had minimal impact on model performance. For example, if OTU 367 was
permuted across the samples in the decision tree model, the median AUROC
dropped from 0.601 to 0.525. In contrast, if the same OTU was permuted
in the random forest model, the AUROC only dropped from 0.695 to 0.680
which indicated high degree of collinearity in the dataset. Permutation
analysis allowed us to gauge the importance of each OTU in non-linear
models and partially account for collinearity by grouping correlated
OTUs to determine their impact as a group.

@@

\textbf{Figure 2. Generalization and classification performance of ML
models using AUROC values of all cross validation and testing
performances.} The median AUROC for diagnosing individuals with SRN
using bacterial abundances was higher than chance (depicted by
horizontal line at 0.50) for all the ML models. The predictive
performance of random forest model was higher than other ML models,
though not significantly (p=0.5). L2-regularized logistic regression,
XGBoost, L2-regularized SVM with linear and radial basis function kernel
performances were not significantly different from one another. The
boxplot shows quartiles at the box ends and the median as the horizontal
line in the box. The whiskers show the farthest points that were not
outliers. Outliers were defined as those data points that are not within
1.5 times the interquartile ranges. \newpage
\includegraphics[height=17cm, width=12cm]{Figure_3.png}

@@

\makecell[l]{Source \\ of data} & \makecell[l]{Input data are not appropriate \\ for intended application (e.g. input \\ data provided only from \\ patients with carcinomas but \\ model is expected to predict \\ advanced adenomas.)} & \makecell[l]{Input data are generated \\ according to study design and \\ they are appropriate for intended \\ application.} & \makecell[l]{Input data are available \\ to predict intended outcomes \\ and will be available \\ in the future (e.g. same OTU \\ assignments for new fecal \\ samples)} \\ \hline

\makecell[l]{Study \\ cohort} & \makecell[l]{Test data resampled to \\ remove class imbalance. \\ (e.g. test data only includes \\ patients with carcinomas but \\ model is trained to predict \\ healthy colons as well.)}  & \makecell[l]{Test data are reflective \\ of the population which \\ the model will eventually \\ be applied to.} & \makecell[l]{Model tested on multiple \\ cohorts with different \\ class balances.} \\ \hline

\makecell[l]{Model \\ selection} & \makecell[l]{No justification for \\ classification method.} & \makecell[l]{Model choice is justified \\ for intended application.} & \makecell[l]{Different modeling \\ choices tested.} \\ \hline

\makecell[l]{Model \\ development} & \makecell[l]{No hyperparameter tuning.} & \makecell[l]{Model is tested across multiple \\ hyperparameter settings.} & \makecell[l]{Hyperparameter grid search \\ performed by cross-validation \\ on the training set.} \\ \hline

\makecell[l]{Model \\ evaluation} & \makecell[l]{Performance reported on the \\data used to train the model.} & \makecell[l]{Performance reported on \\ held-out test data.} & \makecell[l]{Performance reported on \\ multiple held-out test data.} \\ \hline

\makecell[l]{Evaluation \\ metrics} & \makecell[l]{Reported a single performance \\ metric that is not appropriate for \\ intended application  (e.g. when \\ predicting rare outcome, \\ accuracy metric is not \\ reliable).} & \makecell[l]{Reported a single performance \\ metric that is appropriate \\ for intended application with \\ confidence intervals.} & \makecell[l]{Reported multiple metrics \\ with confidence intervals.} \\ \hline

\makecell[l]{Model \\ interpretation} & \makecell[l]{No model interpretation.} & \makecell[l]{Follow-up analyses to \\ determine what is driving \\ model performance.} & \makecell[l]{Hypotheses on important \\ features are generated \\ and tested.} \\ \hline

\makecell[l]{Model \\ specification} & \makecell[l]{Only predictive performance \\ reported.} & \makecell[l]{Model described \\ (e.g. feature weights).} & \makecell[l]{How to use the model \\ is explained.} \\ \hline
\end{tabular}\newpage
